---
title: "Using the Illinois Report Card Data to Teach Statistics"
subtitle: "MMC Conference of Workshops"
date: "2/1/2020"
author: | 
        | Frank Briody 
        | Prospect High School
        | frankbriody@gmail.com
geometry: "left=.5in,right=.5in,top=.5in,bottom=1.0in"
fontsize: 10pt
output: 
  pdf_document: 
    number_sections: yes
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=4.5, fig.height=3) 
options(htmltools.dir.version = FALSE)
#install.packages("ggalt")

library(tidyverse)
library(kableExtra)
library(scales)
library(DT)
library(ggalt) #for dumbbell plot

#theme_set(text = element_text(size=20))
theme_update(text = element_text(size=20))

phs_value <- function(unk) {
  x <- rc17 %>% 
    filter(SCHOOL_ID == "050162140170005")
  x[unk]
}

rc17 <- read_fwf("~/Documents/rc17.txt", 
         fwf_cols(SCHOOL_ID	= c(1,15),
SCHOOL_TYPE_CODE=c(17,17),
SCHOOL_NAME=c(20,52),
DISTRICT_NAME=c(54,86),
CITY=c(88,104),
COUNTY=c(106,116),
DISTRICT_TYPE_CODE=c(118,118),
DISTRICT_SIZE_CODE=c(120,120),
DISTRICT_TYPE_NAME=c(122,132),
DISTRICT_SIZE_NAME=c(134,139),
SCHOOL_TYPE_NAME=c(141,151),
GRADES_IN_SCHOOL=c(153,183),
SCHOOL_WHITE_per=c(185,189),
SCHOOL_BLACK_per=c(191,195),
SCHOOL_HISPANIC_per=c(197,201),
SCHOOL_ASIAN_per=c(203,207),
SCHOOL_NATIVE_HAWAIIAN_OR_OTHER_PACIFIC_ISLANDER_per=c(209,213),
SCHOOL_NATIVE_AMERICAN_per=c(215,219),
SCHOOL_TWO_OR_MORE_RACES_per=c(221,225),
SCHOOL_TOTAL_ENROLLMENT=c(227,231),
LEP_SCHOOL_per=c(385,389),
IEP_SCHOOL_per=c(409,413),
LOWINCOME_SCHOOL_per=c(433,437),
HOMELESS_SCHOOL_per=c(457,461),
TOTAL_SCHOOL_DAYS_SCHOOL=c(481,483),
PARENTAL_INVOLVEMENT_SCHOOL_per=c(497,501),
ATTENDANCE_RATE_SCHOOL_perALL=c(521,525),
ATTENDANCE_RATE_SCHOOL_perMALE=c(545,549),
ATTENDANCE_RATE_SCHOOL_perFEMALE=c(569,573),
ATTENDANCE_RATE_SCHOOL_perWHITE=c(593,597),
ATTENDANCE_RATE_SCHOOL_perBLACK=c(617,621),
ATTENDANCE_RATE_SCHOOL_perHISPANIC=c(641,645),
ATTENDANCE_RATE_SCHOOL_perASIAN=c(665,669),
ATTENDANCE_RATE_SCHOOL_perNATIVE_HAWAIIAN_AND_OTHERS=c(689,693),
ATTENDANCE_RATE_SCHOOL_perNATIVE_AMERICAN=c(713,717),
ATTENDANCE_RATE_SCHOOL_perTWO_OR_MORE_RACES=c(737,741),
ATTENDANCE_RATE_SCHOOL_perLEP=c(761,765),
ATTENDANCE_RATE_SCHOOL_perMIGRANT=c(785,789),
ATTENDANCE_RATE_SCHOOL_perIEP=c(809,813),
ATTENDANCE_RATE_SCHOOL_perLOW_INCOME=c(833,837),
STUDENT_MOBILITY_RATE_SCHOOL_perALL=c(857,861),
STUDENT_MOBILITY_RATE_SCHOOL_perMALE=c(881,885),
STUDENT_MOBILITY_RATE_SCHOOL_perFEMALE=c(905,909),
STUDENT_MOBILITY_RATE_SCHOOL_perWHITE=c(929,933),
STUDENT_MOBILITY_RATE_SCHOOL_perBLACK=c(953,957),
STUDENT_MOBILITY_RATE_SCHOOL_perHISPANIC=c(977,981),
STUDENT_MOBILITY_RATE_SCHOOL_perASIAN=c(1001,1005),
STUDENT_MOBILITY_RATE_SCHOOL_perNATIVE_HAWAIIAN_AND_OTHERS=c(1025,1029),
STUDENT_MOBILITY_RATE_SCHOOL_perNATIVE_AMERICAN=c(1049,1053),
STUDENT_MOBILITY_RATE_SCHOOL_perMULTIRACIAL=c(1073,1077),
STUDENT_MOBILITY_RATE_SCHOOL_perLEP=c(1097,1101),
STUDENT_MOBILITY_RATE_SCHOOL_perIEP=c(1121,1125),
STUDENT_MOBILITY_RATE_SCHOOL_perLOW_INCOME=c(1145,1149),
CHRONIC_TRUANTS_num_SCHOOL=c(1169,1173),
CHRONIC_TRUANTS_RATE_SCHOOL_per=c(1196,1200),
DROPOUT_RATE_SCHOOL_perALL=c(1220,1224),
DROPOUT_RATE_SCHOOL_perMALE=c(1244,1248),
DROPOUT_RATE_SCHOOL_perFEMALE=c(1268,1272),
DROPOUT_RATE_SCHOOL_perWHITE=c(1292,1296),
DROPOUT_RATE_SCHOOL_perBLACK=c(1316,1320),
DROPOUT_RATE_SCHOOL_perHISPANIC=c(1340,1344),
DROPOUT_RATE_SCHOOL_perASIAN=c(1364,1368),
DROPOUT_RATE_SCHOOL_perNATIVE_HAWAIIAN_AND_OTHERS=c(1388,1392),
DROPOUT_RATE_SCHOOL_perNATIVE_AMERICAN=c(1412,1416),
DROPOUT_RATE_SCHOOL_perMULTIRACIAL=c(1436,1440),
DROPOUT_RATE_SCHOOL_perLEP=c(1460,1464),
DROPOUT_RATE_SCHOOL_perMIGRANT=c(1484,1488),
DROPOUT_RATE_SCHOOL_perIEP=c(1508,1512),
DROPOUT_RATE_SCHOOL_perLOW_INCOME=c(1532,1536),
HS_4YEAR_GRAD_RATE_SCHOOL_perALL=c(1556,1560),
HS_4YEAR_GRAD_RATE_SCHOOL_perMALE=c(1580,1584),
HS_4YEAR_GRAD_RATE_SCHOOL_perFEMALE=c(1604,1608),
HS_4YEAR_GRAD_RATE_SCHOOL_perWHITE=c(1628,1632),
HS_4YEAR_GRAD_RATE_SCHOOL_perBLACK=c(1652,1656),
HS_4YEAR_GRAD_RATE_SCHOOL_perHISPANIC=c(1676,1680),
HS_4YEAR_GRAD_RATE_SCHOOL_perASIAN=c(1700,1704),
HS_4YEAR_GRAD_RATE_SCHOOL_perNATIVE_HAWAIIAN_AND_OTHERS=c(1724,1728),
HS_4YEAR_GRAD_RATE_SCHOOL_perNATIVE_AMERICAN=c(1748,1752),
HS_4YEAR_GRAD_RATE_SCHOOL_perMULTIRACIAL=c(1772,1776),
HS_4YEAR_GRAD_RATE_SCHOOL_perLEP=c(1796,1800),
HS_4YEAR_GRAD_RATE_SCHOOL_perMIGRANT=c(1820,1824),
HS_4YEAR_GRAD_RATE_SCHOOL_perIEP=c(1844,1848),
HS_4YEAR_GRAD_RATE_SCHOOL_perLOW_INCOME=c(1868,1872),
ACT_COMP_SCHOOL=c(2564,2567),
ACT_ENGL_SCHOOL_SCORE=c(2584,2587),
ACT_MATH_SCHOOL_SCORE=c(2604,2607),
ACT_READ_SCHOOL_SCORE=c(2624,2627),
ACT_SCIE_SCHOOL_SCORE=c(2644,2647),
per_READY_FOR_COLLEGE_COURSE_WORK_SCHOOL=c(2664,2668),
per_STUDENTS_MET_ENGLISH_BENCHMARK_SCHOOL=c(2688,2692),
per_STUDENTS_MET_MATH_BENCHMARK_SCHOOL=c(2712,2716),
per_STUDENTS_MET_READ_BENCHMARK_SCHOOL=c(2736,2740),
per_STUDENTS_MET_SCI_BENCHMARK_SCHOOL=c(2760,2764),
per_STUDENTS_MET_ALL_4_BENCHMARKS_SCHOOL=c(2784,2788),
AVG_CLASS_SIZE_SCHOOL_HS=c(2988,2991),
OVERALL_AVERAGE_CLASS_SIZE_SCHOOL=c(3008,3011),
per_WHITE_TEACHER_DISTRICT=c(3220,3224),
per_HISPANIC_TEACHER_DISTRICT=c(3244,3248),
per_ASIAN_TEACHER_DISTRICT=c(3256,3260),
per_NATIVE_HAWAIIAN_OR_OTHER_PACIFIC_ISLANDER_TEACHER_DISTRICT=c(3268,3272),
per_NATIVE_AMERICAN_TEACHER_DISTRICT=c(3280,3284),
per_TWO_OR_MORE_RACES_TEACHER_DISTRICT=c(3292,3296),
per_UNKNOWN_RACE_TEACHER_DISTRICT=c(3304,3308),
per_MALE_TEACHER_DISTRICT=c(3316,3320),
per_FEMALE_TEACHER_DISTRICT=c(3328,3332),
TOTAL_TEACH_FTE_COUNT_DISTRICT=c(3340,3346),
AVG_TEACH_EXP_DISTRICT=c(3356,3360),
AVG_TEACH_EXP_DISTRICT_HIGH_POVERTY=c(3368,3372),
AVG_TEACH_EXP_DISTRICT_LOW_POVERTY=c(3374,3378),
BACH_DEGREE_DISTRICT=c(3404,3408),
BACH_DEGREE_DISTRICT_HIGH_POVERTY=c(3416,3420),
BACH_DEGREE_DISTRICT_LOW_POVERTY=c(3422,3426),
MAplus_DEGREE_DISTRICT=c(3452,3456),
MAplus_DEGREE_DISTRICT_HIGH_POVERTY=c(3464,3468),
MAplus_DEGREE_DISTRICT_LOW_POVERTY=c(3470,3474),
PUPILTEACHER_RATIO_HS_DISTRICT=c(3524,3528),
PUPILCERTIFIED_STAFF_RATIO_DISTRICT=c(3548,3552),
PUPILADMIN_RATIO_DISTRICT=c(3572,3576),
AVG_TEACHER_SALARY_DISTRICT=c(3596,3603),
AVG_ADMIN_SALARY_DISTRICT=c(3632,3639),
TEACHER_RETENTION_RATE_SCHOOL=c(3668,3672),
num_OF_PRINCIPAL_TURNOVER_WITHIN_6_YEARS_SCHOOL=c(3692,3695),
LOCAL_PROPERTY_TAXES_DISTRICT_DOLLAR201516=c(3712,3726),
LOCAL_PROPERTY_TAXES_DISTRICT_per201516=c(3728,3732),
OTHER_LOCAL_FUNDING_DISTRICT_DOLLAR201516=c(3740,3754),
OTHER_LOCAL_FUNDING_DISTRICT_per201516=c(3756,3760),
GENERAL_STATE_AID_DISTRICT_DOLLAR201516=c(3768,3782),
GENERAL_STATE_AID_DISTRICT_per201516=c(3784,3788),
OTHER_STATE_FUNDING_DISTRICT_DOLLAR201516=c(3796,3810),
OTHER_STATE_FUNDING_DISTRICT_per201516=c(3812,3816),
FEDERAL_FUNDING_DISTRICT_DOLLAR201516=c(3824,3838),
FEDERAL_FUNDING_DISTRICT_per201516=c(3840,3844),
TOTAL_REVENUE_SOURCE201516=c(3852,3866),
EAV_PER_PUPIL_DISTRICT2014=c(3868,3877),
TOT_SCHL_TAX_RATE_PER_100_DISTRICT2014=c(3901,3905),
INSTRUCT_EXPEND_PER_PUPIL_DISTRICT201516=c(3919,3926),
OPER_EXPEND_PP_DISTRICT201516=c(3955,3962),
INSTRUCTION__DOLLARS201516=c(3991,4005),
INSTRUCTION__DISTRICT_per201516=c(4007,4011),
GENERAL_ADMIN__DOLLARS201516=c(4019,4033),
GENERAL_ADMIN__DISTRICT_per201516=c(4035,4039),
SUPPORTING_SERVICES__DOLLARS201516=c(4047,4061),
SUPPORTING_SERVICES__DISTRICT_per201516=c(4063,4067),
OTHER_EXPENDITURES__DOLLARS201516=c(4075,4089),
OTHER_EXPENDITURES__DISTRICT_per201516=c(4091,4095),
EDUCATION_FUND_DOLLARS201516=c(4103,4117),
EDUCATION_FUND_DISTRICT_per201516=c(4119,4123),
OPERATIONS_B_M_DOLLARS201516=c(4131,4145),
OPERATIONS_B_M_DISTRICT_per201516=c(4147,4151),
TRANSPORTATION_DOLLARS201516=c(4159,4173),
TRANSPORTATION_DISTRICT_per201516=c(4175,4179),
DEBT_SERVICE_DOLLARS201516=c(4187,4201),
DEBT_SERVICE_DISTRICT_per201516=c(4203,4207),
TORT_DOLLARS201516=c(4215,4229),
TORT_DISTRICT_per201516=c(4231,4235),
MUNIC_RET_SOC_SECURITY_DOLLARS201516=c(4243,4257),
MUNIC_RET_SOC_SECURITY_DISTRICT_per201516=c(4259,4263),
FIRE_PREV_SAFETY_DOLLARS201516=c(4271,4285),
FIRE_PREV_SAFETY_DISTRICT_per201516=c(4287,4291),
CAPITAL_PROJECTS_FUND_DOLLARS201516=c(4299,4313),
CAPITAL_PROJECTS_FUND_DISTRICT_per201516=c(4315,4319),
TOTAL_EXPENDITURES201516=c(4327,4341),
LEGISLATIVE_SCHOOL=c(4347,4348),
LEGISLATIVE_DISTRICT=c(4350,4351),
per_GRADUATES_ENROLLED_IN_COLLEGE_WITHIN_16_MONTHS_SCHOOL=c(5457,5461),
per_GRADUATES_ENROLLED_IN_COLLEGE_WITHIN_12_MONTHS_SCHOOL=c(5481,5485),
per_FRESHMAN_ON_TRACK_SCHOOL=c(5505,5509),
num_of_CTE_ENROLLMENT_SCHOOL=c(5553,5557),
num_OF_GRADE_10_STUDENTS_TOOK_ADVANCED_COURSES_SCHOOLALL=c(5585,5589),
num_OF_GRADE_11_STUDENTS_TOOK_ADVANCED_COURSES_SCHOOLALL=c(5611,5615),
num_OF_GRADE_12_STUDENTS_TOOK_ADVANCED_COURSES_SCHOOLALL=c(5637,5641),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLALL=c(5663,5667),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLWHITE=c(5689,5693),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLBLACK=c(5715,5719),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLHISPANIC=c(5741,5745),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLASIAN=c(5767,5771),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNATIVE_HAWAIIAN_AND_OTHERS=c(5793,5797),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNATIVE_AMERICAN=c(5819,5823),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLTWO_OR_MORE_RACES=c(5845,5849),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLLEP=c(5871,5875),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNONLEP=c(5897,5901),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLIEP=c(5923,5927),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNONIEP=c(5949,5953),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLLOW_INCOME=c(5975,5979),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNONLOW_INCOME=c(6001,6005),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLALL=c(6027,6031),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLWHITE=c(6053,6057),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLBLACK=c(6079,6083),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLHISPANIC=c(6105,6109),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLASIAN=c(6131,6135),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNATIVE_HAWAIIAN_AND_OTHERS=c(6157,6161),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNATIVE_AMERICAN=c(6183,6187),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLTWO_OR_MORE_RACES=c(6209,6213),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLLEP=c(6235,6239),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNONLEP=c(6261,6265),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLIEP=c(6287,6291),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNONIEP=c(6313,6317),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLLOW_INCOME=c(6339,6343),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNONLOW_INCOME=c(6365,6369),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLALL=c(6391,6395),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLWHITE=c(6417,6421),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLBLACK=c(6443,6447),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLHISPANIC=c(6469,6473),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLASIAN=c(6495,6499),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNATIVE_HAWAIIAN_AND_OTHERS=c(6521,6525),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNATIVE_AMERICAN=c(6547,6551),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLTWO_OR_MORE_RACES=c(6573,6577),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLLEP=c(6599,6603),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNONLEP=c(6625,6629),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLIEP=c(6651,6655),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNONIEP=c(6677,6681),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLLOW_INCOME=c(6703,6707),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_COURSES_SCHOOLNONLOW_INCOME=c(6729,6733),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLALL=c(6755,6759),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLWHITE=c(6781,6785),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLBLACK=c(6807,6811),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLHISPANIC=c(6833,6837),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLASIAN=c(6859,6863),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNATIVE_HAWAIIAN_AND_OTHERS=c(6885,6889),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNATIVE_AMERICAN=c(6911,6915),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLTWO_OR_MORE_RACES=c(6937,6941),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLLEP=c(6963,6967),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNONLEP=c(6989,6993),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLIEP=c(7015,7019),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNONIEP=c(7041,7045),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLLOW_INCOME=c(7067,7071),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNONLOW_INCOME=c(7093,7097),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLALL=c(7119,7123),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLWHITE=c(7145,7149),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLBLACK=c(7171,7175),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLHISPANIC=c(7197,7201),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLASIAN=c(7223,7227),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNATIVE_HAWAIIAN_AND_OTHERS=c(7249,7253),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNATIVE_AMERICAN=c(7275,7279),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLTWO_OR_MORE_RACES=c(7301,7305),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLLEP=c(7327,7331),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNONLEP=c(7353,7357),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLIEP=c(7379,7383),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNONIEP=c(7405,7409),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLLOW_INCOME=c(7431,7435),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNONLOW_INCOME=c(7457,7461),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLALL=c(7483,7487),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLWHITE=c(7509,7513),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLBLACK=c(7535,7539),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLHISPANIC=c(7561,7565),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLASIAN=c(7587,7591),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNATIVE_HAWAIIAN_AND_OTHERS=c(7613,7617),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNATIVE_AMERICAN=c(7639,7643),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLTWO_OR_MORE_RACES=c(7665,7669),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLLEP=c(7691,7695),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNONLEP=c(7717,7721),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLIEP=c(7743,7747),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNONIEP=c(7769,7773),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLLOW_INCOME=c(7795,7799),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_IB_COURSES_SCHOOLNONLOW_INCOME=c(7821,7825),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLALL=c(7847,7851),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLWHITE=c(7873,7877),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLBLACK=c(7899,7903),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLHISPANIC=c(7925,7929),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLASIAN=c(7951,7955),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNATIVE_HAWAIIAN_AND_OTHERS=c(7977,7981),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNATIVE_AMERICAN=c(8003,8007),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLTWO_OR_MORE_RACES=c(8029,8033),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLLEP=c(8055,8059),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNONLEP=c(8081,8085),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLIEP=c(8107,8111),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNONIEP=c(8133,8137),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLLOW_INCOME=c(8159,8163),
num_OF_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNONLOW_INCOME=c(8185,8189),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLALL=c(8211,8215),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLWHITE=c(8237,8241),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLBLACK=c(8263,8267),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLHISPANIC=c(8289,8293),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLASIAN=c(8315,8319),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNATIVE_HAWAIIAN_AND_OTHERS=c(8341,8345),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNATIVE_AMERICAN=c(8367,8371),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLTWO_OR_MORE_RACES=c(8393,8397),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLLEP=c(8419,8423),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNONLEP=c(8445,8449),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLIEP=c(8471,8475),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNONIEP=c(8497,8501),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLLOW_INCOME=c(8523,8527),
num_OF_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNONLOW_INCOME=c(8549,8553),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLALL=c(8575,8579),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLWHITE=c(8601,8605),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLBLACK=c(8627,8631),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLHISPANIC=c(8653,8657),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLASIAN=c(8679,8683),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNATIVE_HAWAIIAN_AND_OTHERS=c(8705,8709),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNATIVE_AMERICAN=c(8731,8735),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLTWO_OR_MORE_RACES=c(8757,8761),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLLEP=c(8783,8787),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNONLEP=c(8809,8813),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLIEP=c(8835,8839),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNONIEP=c(8861,8865),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLLOW_INCOME=c(8887,8891),
num_OF_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_DUAL_CREDIT_COURSES_SCHOOLNONLOW_INCOME=c(8913,8917),
num_of_AP_EXAMS_TAKEN_BY_GRADE_10_STUDENTS_SCHOOL=c(8939,8943),
num_of_AP_EXAMS_PASSED_BY_GRADE_10_STUDENTS_SCHOOL=c(8961,8965),
num_of_GRADE_10_STUDENTS_TOOK_ONE_OR_MORE_AP_EXAMS_SCHOOL=c(8983,8987),
num_of_GRADE_10_STUDENTS_PASSED_ONE_OR_MORE_AP_EXAMS_SCHOOL=c(9005,9009),
num_of_AP_EXAMS_TAKEN_BY_GRADE_11_STUDENTS_SCHOOL=c(9027,9031),
num_of_AP_EXAMS_PASSED_BY_GRADE_11_STUDENTS_SCHOOL=c(9049,9053),
num_of_GRADE_11_STUDENTS_TOOK_ONE_OR_MORE_AP_EXAMS_SCHOOL=c(9071,9075),
num_of_GRADE_11_STUDENTS_PASSED_ONE_OR_MORE_AP_EXAMS_SCHOOL=c(9093,9097),
num_of_AP_EXAMS_TAKEN_BY_GRADE_12_STUDENTS_SCHOOL=c(9115,9119),
num_of_AP_EXAMS_PASSED_BY_GRADE_12_STUDENTS_SCHOOL=c(9137,9141),
num_of_GRADE_12_STUDENTS_TOOK_ONE_OR_MORE_AP_EXAMS_SCHOOL=c(9159,9163),
num_of_GRADE_12_STUDENTS_PASSED_ONE_OR_MORE_AP_EXAMS_SCHOOL=c(9181,9185),
per_OF_GRADUATES_ENROLLED_IN_COMMUNITY_COLLEGE_REMEDIAL_COURSES_SCHOOLALL=c(9203,9207),
STUDENT_RESPONSE_RATE_SCHOOL_DATA=c(9227,9231),
TEACHER_RESPONSE_RATE_SCHOOL_DATA=c(9233,9237),
AVERGE_NUMBER_OF_DAYS_OF_PHYSICAL_EDUCATION_PER_WEEK_PER_STUDENT_SCHOOL=c(9263,9263),
TEACHER_ATTENDANCE_RATE_SCHOOLALL=c(9273,9277),
per_of_TEACHERS_RATED_AS_PROFICIENT_OR_EXCELLENT_DISTRICT=c(9297,9301),
num_OF_SCHOOLS_IN_THE_DISTRICT_AND_OTHER_ADMINISTRATIVE_ENTITIES_FOR_REPORT_CARD=c(9309,9311)),
progress = TRUE)

rc17$CHRONIC_TRUANTS_num_SCHOOL <- as.numeric(gsub('[$,]', '', rc17$CHRONIC_TRUANTS_num_SCHOOL))
rc17$num_of_AP_EXAMS_TAKEN_BY_GRADE_10_STUDENTS_SCHOOL <- as.numeric(gsub('[$,]', '', rc17$num_of_AP_EXAMS_TAKEN_BY_GRADE_10_STUDENTS_SCHOOL))
rc17$num_of_AP_EXAMS_PASSED_BY_GRADE_11_STUDENTS_SCHOOL <- as.numeric(gsub('[$,]', '', rc17$num_of_AP_EXAMS_PASSED_BY_GRADE_11_STUDENTS_SCHOOL))
rc17$AVG_TEACHER_SALARY_DISTRICT <- as.numeric(gsub('[$,]', '',rc17$AVG_TEACHER_SALARY_DISTRICT))
rc17$AVG_ADMIN_SALARY_DISTRICT <- as.numeric(gsub('[$,]', '',rc17$AVG_ADMIN_SALARY_DISTRICT))
rc17$TOTAL_REVENUE_SOURCE201516 <- as.numeric(gsub('[$,]', '',rc17$TOTAL_REVENUE_SOURCE201516))
rc17$TOTAL_EXPENDITURES201516 <- as.numeric(gsub('[$,]', '',rc17$TOTAL_EXPENDITURES201516))
rc17$INSTRUCT_EXPEND_PER_PUPIL_DISTRICT201516 <- as.numeric(gsub('[$,]', '',rc17$INSTRUCT_EXPEND_PER_PUPIL_DISTRICT201516))
rc17$EAV_PER_PUPIL_DISTRICT2014 <- as.numeric(gsub('[$,]', '',rc17$EAV_PER_PUPIL_DISTRICT2014))

sixco <- rc17 %>% 
  filter(COUNTY %in% c("Cook", "Lake", "Will", "Kane", "McHenry", "Dupage"))

sixco_hs <- sixco %>% 
  filter(SCHOOL_TYPE_NAME == "HIGH SCHOOL")
```

\newpage

# Variables
The ISBE raw data file *rx17.txt* contains 1,471 variables. The variable definitions are in the Excel file *RC17_layout.xlsx* and have been categorized into the groups shown below. The first number represents available variables in each group while the second is the number actually imported into the processed data file. The import script produces 316 variables from 20 of the 21 categories for all 3,796 Illinois public schools. (None of the NAEP variables were imported.) Usable files will be discussed in section \ref{import}.  

\rule[1ex]{\textwidth}{.1pt}

\begin{minipage}[t]{0.5\textwidth}
School information (13 variables;12 imported)\\
Student demographics (396;71)\\
ACT (44;11)\\
Instructional setting (92;2)\\
Teacher and admin statistics (78;26)\\
District financial (67;40)\\
Region and legislative (3;2)\\
National Assmnt. of Educ. Progress (NAEP) (184;0)\\
College and Career readiness (16;3)
CTE (4;1)\\
Advanced coursework (12;3)\\
\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
AP courses (168;42)\\
IB courses (168;42)\\
Dual credit (168;42)\\
AP exams (36;12)\\
Post secondary remediation (4;1)\\
Response rate (5E survey) (4;2)\\
Health and wellness (3;1)\\
Teacher Attendance (4;1)\\
Teacher Evaluation (2;1)\\
School District Count (3;1)
\end{minipage}

\rule[1ex]{\textwidth}{.1pt}

# Descriptive Statistics via State Demographics
## Categorical Count (Raw)

```{r}
school_type <- rc17 %>% 
  count(SCHOOL_TYPE_NAME, sort = TRUE) %>% 
  mutate(rel_freq = n/sum(n))
school_type
```



## Categorical Count (Formatted)

```{r}
kable(school_type) %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

## Categorical Plot

```{r}
ggplot(rc17, aes(x=factor(SCHOOL_TYPE_NAME)))+
  geom_bar(fill="#2F4E6F")+
  labs(title = "Type of School", x = "School Type", subtitle = "All Illinois Schools (2017)") +
  theme_minimal()
```

## Categorical Analysis I 
Write a short analysis for the types of schools in the state of Illinois.

\vspace{75pt}


## Categorical Analysis II 

```{r fig.width=6, fig.height=4}
rc17 %>% 
filter(COUNTY == "Dupage" | COUNTY == "Will" | COUNTY == "Kane" | 
         COUNTY == "Lake" | COUNTY == "Cook"| COUNTY == "McHenry") %>% 
ggplot(aes(x=factor(SCHOOL_TYPE_NAME), y = (..count..)/sum(..count..))) +
  geom_bar(fill="#2F4E6F")+
  facet_wrap(~COUNTY, nrow = 2) +
  labs(title = "Type of School by County", 
       x = "School Type", 
       y = "proportion",
       subtitle = "Six Counties in the Chicago Metropolitan Region (2017)") +
  theme_minimal()  +
  theme(axis.text.x = element_text(angle = 60, vjust = 0.5))
```

Write a short analysis for the types of schools in the six county region.

\vspace{75pt}


# Software
## R and RStudio
R is open source, free, industry-standard statistical analysis software that was first introduced in 1993. (R is an adaptation of the S languge that was invented at Bell Labs in 1976.) RStudio, which arrived in 2011, is a free development environment for using R. Both R and RStudio have a bit of an initial learing curve, but mastering a few basic commands opens up a world of analysis options. The RStudio installation page [https://rstudio.com/products/rstudio/download/#download] provides instructions for setting up both R and RStudio.  Suggestions for learning R are in the \nameref{sec:Appendix}. 


## Fathom
Key Curriculum Press, the creators of The Geometer's Sketchpad software, sells Fathom ($39 USD) at [https://fathom.concord.org]. Fathom is graphical statistical analysis software which accomplishes most tasks through a drag and drop interface. Its primary audience is teachers and students, but has reached end-of-life status. It can still be purchased and used but no further developments are expected. 

## Others
Excel, JMP, and SAS are other software platforms that can be used for analysis but will not be discussed here. The report card data will be provided in .csv format which can be imported by these programs. 

# Data Import\label{import}
## Data Files

INSERT GRAPHIC HERE

### Super Easy Method
All graphs from this presentation are available for download at [INSERT MMC GITHUB HTML LINK]. ???Look in the appendix section.

### Easy Method
A reasonably sized (2.7MB) data file containing 316 variables for the 2,049 schools in the Chicagoland six county region is avaialble at [GITHUB LINK]. The six counties are Cook, Dupage, Kane, Lake, McHenry and Will. This data file can be imported into RStudio, Fathom, Excel, etc. 

### Some Variables for All Schools 
I used the import script ```import_rc17.txt``` to select 316 variables for all 3,796 Illinois schools. A 5MB data file is available at []. Again, this data file can be imported into RStudio, Fathom, Excel, etc. 

### Starting from Scratch
The original raw data is available on the ISBE Report Card Data Library web page  [https://www.isbe.net/Pages/Illinois-State-Report-Card-Data.aspx]. You will need to download both the fixed width data file (rc17.txt 2.4MB becomes 35.4MB) and the variable definitions (RC17_layout.xlsx). Use and/or modify import_rc17.txt [https://github.com/fbriody/MMC2020] to get a subset of the data into RStudio. 


Subsetting and exporting is a two step process. First, use R to create the subset:
```
sixco <- rc17 %>% 
  filter(COUNTY %in% c("Cook", "Lake", "Will", "Kane", "McHenry", "Dupage"))
```
Then, export:
```
write.csv(sixco,"sixco.csv", row.names = FALSE)
```

The resulting .csv can be imported by another software platform. (Note there is no need to export if you are staying in Studio. Just refer to you r new subset, in this case `sixco`.) You can customize the above command(s) to suit your needs. 


## Processed into Fathom
### Import
csv import attribute names

### Displays
missing values and shift or option
inspecting
adding a third variable
  act v LowInc with school type 

## Importing .csv into Excel
Use Excel to import rc17.csv. *****CHANGE NAME to rc17_fb.csv??*****

# Numeric Summaries (REORDER THESE SECTIONS)
## Lists
### Number of High Schools in the Six County Region
```{r}
sixco %>% 
  filter(SCHOOL_TYPE_NAME == "HIGH SCHOOL") %>% 
  group_by(COUNTY) %>% 
  summarise(count = n())
```

### Single List of Scores
Subsetting based on a criteria produces a single list of scores. 

```{r}
mchenry_act <- rc17 %>% 
  filter(COUNTY == "McHenry", is.na(ACT_COMP_SCHOOL) == FALSE )
mchenry_act$ACT_COMP_SCHOOL
```

Adding ```sort()``` orders the scores. Remove the comma and the ```decreasing``` option to priduce an increasing list. 

```{r}
sort(mchenry_act$ACT_COMP_SCHOOL, decreasing = TRUE)
```

### Lake County ACT Scores (Ordered and Formatted)

```{r}
lake_ACT <- rc17 %>% 
  filter(SCHOOL_TYPE_NAME == "HIGH SCHOOL", COUNTY == "Lake") %>% 
  arrange(desc(ACT_COMP_SCHOOL)) %>% 
  select(COUNTY, SCHOOL_NAME, ACT = ACT_COMP_SCHOOL)
kable(lake_ACT)
```

Create a boxplot for Lake County ACT scores. How could you compare to DuPage county?

## Two-Way Tables

```{r}
district_type <- rc17 %>% 
  filter(COUNTY == "Lake" | COUNTY == "Dupage") %>% 
  group_by(COUNTY)

two_way <- with(district_type, table(DISTRICT_SIZE_NAME, COUNTY))

kable(two_way, caption = "Types of School Districts") %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

## Resistant Measures
```{r echo = FALSE, fig.height=4, fig.width=6}
sixco  %>% 
  filter(SCHOOL_TYPE_NAME == "HIGH SCHOOL") %>% 
  ggplot(mapping = aes(x = COUNTY, y = ACT_COMP_SCHOOL)) +
  geom_boxplot() + 
  theme_minimal() + 
  coord_flip() +
  labs(x = "County", y = "Composite ACT", 
       title = "Composite ACT Scores for High Schools")
```


## Mean vs Median
### Instructional Spending per Pupil by District
#### Numeric Summary
```{r}
rc17 %>% 
  filter(SCHOOL_TYPE_NAME == "HIGH SCHOOL") %>% 
  group_by(DISTRICT_NAME) %>% 
  summarise(spend_stu = mean(INSTRUCT_EXPEND_PER_PUPIL_DISTRICT201516, na.rm = TRUE)) %>% 
  summary()
```
#### Plot

```{r echo = FALSE,  warning = FALSE,   fig.width=6, fig.height=4}
spend_stu_plot <- rc17 %>% 
  filter(SCHOOL_TYPE_NAME == "HIGH SCHOOL") %>% 
  group_by(DISTRICT_NAME) %>% 
  summarise(spend_stu = mean(INSTRUCT_EXPEND_PER_PUPIL_DISTRICT201516, na.rm = TRUE)) %>% 
  ggplot(mapping = aes(x = spend_stu)) +
  geom_histogram(binwidth = 500, color = "Blue", fill = "LightBlue") +
  scale_x_continuous(labels = dollar)

spend_stu_plot +
geom_segment(aes(x=5912, y=62.5, xend=5912, yend=0), 
             arrow = arrow(length = unit(.03, "npc")), 
             color="DarkGreen") +
  geom_segment(aes(x=6571, y=75, xend=6571, yend=0), 
              arrow = arrow(length = unit(.03, "npc")), 
              color="DarkRed") + 
  geom_label(aes(x=5912, y=62.5,  
              label = "Median \n $5,912"), 
              nudge_x = -400) + 
  geom_label(aes(x=6571, y=75,  
              label = "Mean \n $6,571"), 
              nudge_x = 400) + 
  geom_segment(aes(x=12405, y=80, xend=12405, yend=0), 
              arrow = arrow(length = unit(.03, "npc")), 
              color="DarkBlue") + 
  geom_label(aes(x=12405, y=80,  
              label = "PHS \n $12,405"), 
              nudge_x = 400) + 
  theme_minimal()

```

# Single Values
## Finding a School
```{r}
rc17 %>% 
  filter(str_detect(SCHOOL_NAME, "Morton")) %>% 
  select(SCHOOL_ID, SCHOOL_NAME, ACT_COMP_SCHOOL)
```

## Using a Filter
```{r}
prospect <- rc17 %>% 
  filter(str_detect(SCHOOL_NAME, "Prospect High School"))
prospect_act <- prospect$ACT_COMP_SCHOOL
prospect_act
```

## Using a Function
```{r}
phs_value <- function(unk) {
  x <- rc17 %>% 
    filter(SCHOOL_ID == "050162140170005")
  x[unk]
}

phs_value("ACT_COMP_SCHOOL")
```

## Single Values in Fathom



# Correlation and Regression 
## Guess the Correlation
- ACT Composite vs Chronically Truant (#) Guess:___________ Actual:___________
- ACT Composite vs Chronically Truant (%) Guess:___________ Actual:___________
- ACT Composite vs Student Mobility Guess:___________ Actual:___________
- ACT Composite vs Attendance rate (%) Guess:___________ Actual:___________

## Predicting  ACT Scores
Suppose you choose 3 variables (Composite ACT Score, Enrollment and Attendance Rate) for all schools in the Six County region. What question(s) and display(s) would you explore?

\newpage

## Scatterplot Analysis
### Outliers and Influential
```{r fig.width=6, fig.height=4}
ACTvLI <- sixco %>% 
  ggplot(mapping = aes(x = LOWINCOME_SCHOOL_per, y = ACT_COMP_SCHOOL)) +
  geom_point(color="Blue") +
  theme_minimal()
ACTvLI
```

Add some labels using the ```ggrepel``` package. (Also notice the layering of information.)
```{r}
library(ggrepel)
```

```{r fig.width=6, fig.height=4}
ACTvLI + 
  geom_text(aes(label=ifelse(SCHOOL_ID =="050162140170005",
                             as.character(SCHOOL_NAME),'')),hjust=0,vjust=0) +
  geom_text_repel(aes(LOWINCOME_SCHOOL_per, ACT_COMP_SCHOOL, 
                      label = ifelse(ACT_COMP_SCHOOL >25 & 
                                     LOWINCOME_SCHOOL_per>25, 
                                     SCHOOL_NAME, "")), size = 3)
```


### Adding a Third Variable
```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
sixco %>% ggplot(mapping = aes(x= ATTENDANCE_RATE_SCHOOL_perALL, y = ACT_COMP_SCHOOL)) +
  geom_point(aes(alpha = sixco$SCHOOL_TOTAL_ENROLLMENT)) +
  labs(alpha = "Enrollment", x = "Attendance Rate", y = "ACT Composite", title = "Predicting ACT from Attendance", subtitle = "Six County High Schools") +
  theme(legend.position = c(.165, .75), text = element_text(size=10))
```

### Regression in Fathom
Force a variable to be numeric (option) or categorical (shift). 
Adding a Third Variable: Are charter schools different?

## Regression Output
### All Schools
```{r}
summary(lm(sixco$ACT_COMP_SCHOOL~sixco$ATTENDANCE_RATE_SCHOOL_perALL))
```

### Influential Removed
 
```{r}
sixco_removed <- sixco %>% 
  filter(ATTENDANCE_RATE_SCHOOL_perALL>50) 

  summary(lm(sixco_removed$ACT_COMP_SCHOOL~sixco_removed$ATTENDANCE_RATE_SCHOOL_perALL))
```

### Both Models


```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
sixco %>% 
  filter(ATTENDANCE_RATE_SCHOOL_perALL>50) %>% 
  ggplot(mapping = aes(x= ATTENDANCE_RATE_SCHOOL_perALL, y = ACT_COMP_SCHOOL, alpha = SCHOOL_TOTAL_ENROLLMENT)) +
  geom_point() + 
  geom_smooth(method = lm, na.rm = TRUE, se = FALSE) + 
  geom_abline(intercept = -11.07428, slope = 0.34033, color = "Red") + 
  theme(legend.position = "none",text = element_text(size=10))

```

### Residual Plot

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
ggplot(lm(sixco$ACT_COMP_SCHOOL~sixco$ATTENDANCE_RATE_SCHOOL_perALL)) + 
  geom_point(mapping = aes(x = .fitted, y = .resid)) + 
  geom_hline(yintercept = 0, color = "Blue") +
  labs(x = "Predicted ACT Composite", 
       y = "Residual", 
       title = "Residual Plot for Predicting School ACT from School Attendance Rate (%)") + 
  theme(text = element_text(size = 10))

```

# Random Selection and Simulation
## Rolling a die

```{r}
set.seed(2020)
one_die <- sample(1:6, 10, replace = TRUE)
one_die
```

## Random Selection
```{r}
four_schools <-  sample_n(rc17, 4) 
four_schools[c("SCHOOL_NAME", "SCHOOL_TOTAL_ENROLLMENT")]
mean(four_schools$SCHOOL_TOTAL_ENROLLMENT)
```

## Stratified Sample
```{r}
strat_samp <- sixco %>% 
  filter(SCHOOL_TYPE_NAME == "HIGH SCHOOL") %>% 
  group_by(COUNTY) %>% 
  sample_n(3)
strat_samp[c("SCHOOL_NAME", "COUNTY", "SCHOOL_TOTAL_ENROLLMENT")]
```

## Confidence Interval Simulation
INCLUDE SD IN TABLE SO CAN CALCULATE INTERVAL BY HAND
```{r}
rand_samp <- function(samples, vari, samp_size) {
  sixco_hs_nona <- sixco_hs[!is.na(sixco_hs[vari]), ] #remove schools with no value
  a <- matrix(ncol = 6, nrow = samples)
  for (k in 1:samples){
    dat_fra <- sample_n(sixco_hs_nona, samp_size)
    t_star <- qt(.975, df = samp_size - 1)
    x_bar <- mean(dat_fra[[vari]])
    stan_dev <- sd(dat_fra[[vari]])
    lower_b <- x_bar  - t_star*stan_dev/(samp_size)**.5
    upper_b <- x_bar + t_star*stan_dev/(samp_size)**.5
  v1 <- k
  v2 <- x_bar
  v3 <- lower_b
  v4 <- upper_b
  v5 <- mean(sixco_hs[[vari]], na.rm = TRUE)
  v6 <- samp_size
  a[k,] <- c(v1, v2, v3, v4, v5, v6)}
  colnames(a) <- c("Sample", "Mean", "L_Bound", "U_Bound", "Parameter", "Sample_Size")
  return(a)
  }
```

sum(dat_fra[[vari]] !="", na.rm = TRUE)

```{r}
confid_ints <- as_tibble(rand_samp(29, "ACT_COMP_SCHOOL", 25))
confid_ints
```


```{r fig.width=6, fig.height=4}
ggplot(confid_ints, mapping = aes(x=L_Bound, xend = U_Bound, y = Sample)) + 
  geom_point(aes(x=Mean, y=Sample)) +
    geom_vline(xintercept = mean(confid_ints$Parameter), 
               linetype="dashed", 
                color = "grey", 
               size=1) +
    geom_dumbbell(size_xend=0,size_x=0,
                color = ifelse(confid_ints$U_Bound < confid_ints$Parameter |
                                confid_ints$L_Bound > confid_ints$Parameter,
                                "red", "black")) + 
    labs(x = "", 
         title = paste(max(confid_ints$Sample), 
                       "Samples of size n =",
                       max(confid_ints$Sample_Size)
                      ),
         subtitle = paste("Parameter = ",round(confid_ints$Parameter,2))
         ) + 
  theme_minimal()
```



## Binomial
A basketball player claims he makes 70% of his free throws. During a recent game he made only 4 of 10. Does this cast doubt on his 70% claim or could making only 4 of 10 happen to a 70% shooter?
```{r}
free_throws <- rbinom(300, 10, .7)
free_throws
sort(free_throws)
table(free_throws)
```
 
# Appendix\label{sec:Appendix}
## Learning More
### R and RStudio
R4DS
### Fathom
### Statistics
DePaul, Udacity, CodeAcademy, DataCamp
Generic Graphs
